---
title: "x) Creating data objects for an R Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{x) Creating data objects for an R Package}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


For overview read https://r-pkgs.org/data.html

## Preliminaries

```{r, eval = F}
library(usethis)
library(here)
```



##  Create  /data-raw

This creates the folder /data-raw and adds a file DATASET.R

```{r, eval = F}
usethis::use_data_raw()
```

## Create data processing scripts

### Create script file

```{r, eval = F}
usethis::use_data_raw(name = "human_gene_lengths")
```

### Get the data

In this case I'm using data from Whitlock and Schulter's 2nd edition of [*Analysis of Biological Data*](https://www.amazon.com/Analysis-Biological-Data-Michael-Whitlock/dp/0981519407), downloaded from the [book's website](https://whitlockschluter.zoology.ubc.ca/) https://whitlockschluter.zoology.ubc.ca/

```{r, eval = F}
human_gene_lengths <- read.csv(url("http://whitlockschluter.zoology.ubc.ca/wp-content/data/chapter04/chap04e1HumanGeneLengths.csv"))

names(human_gene_lengths) <- "gene_length"
```

### Create the .RData file

This converts and R object into a .RData file that can be loaded with data().

If necessary it creates a data/ folder for the package.

```{r, eval = F}
usethis::use_data(human_gene_lengths, 
                  overwrite = TRUE)
```

### Create documentation file for the data

All datasets need a .R file the goes in the R/ folder, along with any .R files that define functions.


```{r, eval = F}
usethis::use_r(name = "human_gene_lengths", open = T)
```

This is an opportunity though to provide full documentation for the dataset.  A minimal helpfile could look like this:

```{r, eval = F}
#' Dataset helpfile header . . .
#'
#' Short description of data . . .
#'
#' @format A data frame with x rows and y column(s)
#' \describe{
#'   \item{column1}{Describe column here . . .}
#'   \item{column2}{Describe column here . . .}
#'  ...
#' }
#' @source \url{http://www.whereyougotthedata.com/"}
#'
"dataset_name"
```

You can also add additional things that appear in R help files such as full citation information and examples.

For other examples see https://github.com/hadley/babynames/blob/master/R/data.R



The function defined below can be used to build this helpfile template automatically.  It is also found in my [biodata](https://github.com/brouwern/biodata) package.

```{r, eval = F}
# Function to build template for dataset helpf file
make_dateset_helpfile <- function(dataset,
                                  dataset_name  = "temp"){
  library(here)

  dataset <- human_gene_lengths
  dataset_name <- "human_gene_lengths"
  to_sink <- paste(dataset_name,"R",sep = ".")
  to_sink_with_dir <- here::here("R",to_sink)

  sink(to_sink_with_dir)
  cat("#' Dataset helpfile header . . .\n")
  cat("#'\n")
  cat("#' Short description of data . . . \n")
  cat("#'\n")
  cat("#' @format A data frame with ", dim(dataset)[1], " rows and ",dim(dataset)[2]," column(s)\n", sep = "")
  cat("#' \\describe{\n", sep = "")

  for(i in 1:ncol(dataset)){
    colname.i <- names(dataset)[i]
    cat("#'   \\item{",colname.i,"}{Describe column ",colname.i, " here . . .}\n",sep = "")
  }

  cat("#' }\n")
  cat("#' @source \\url{http://www.whereyougotthedata.com/}\n")
  cat("#'\n")
  cat("'",dataset_name,"'", sep = "")
  sink()
}


```



