---
title: "y) Building & rebuilding this package"
author: "Nathan Brouwer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{y) Building & rebuilding this package}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval  = F,
  comment = "#>"
)
```

## Introduction

This code is used to build, re-build, and trouble shoot the building of the package. 

These are mostly commands that can be launched from within RStudio via the GUI, except pkgdown::build_site()

## devtools functions

This requires the devtools package.

```{r}
library(devtools)
```

## Vignettes

I build and troubleshoot vignettes by knitting them individually, then when things are running smooth build the package.

### Clean vignettes

Cleans up stray files associated vignettes.

```{r, eval = F}
devtools::clean_vignettes() 
```

### Build vignettes

devtools::build_vignettes() does the same things call
R CMD build 

```{r, eval = F}
devtools::build_vignettes()
```


**NOTE:** The most common error I get here is 

```{r, eval = F}
Error: callr subprocess failed: Duplicate vignette titles.
  Ensure that the %\VignetteIndexEntry lines in the vignette sources
  correspond to the vignette titles.
```


This occurs because there's a pesky bit in the YAML header of Vigenettes

```{r, eval = F}
%\VignetteIndexEntry{...}
```


If you copy and paste the YAML header when you make a new vignette, you need to remember change this.

### Build documentation

document()

```{r, eval = F}
devtools::document(roclets=c('rd', 'collate', 'namespace'))
```

### Load all

```{r, eval = F}
devtools::load_all(".")

```


### Check

```{r, eval = F}
devtools::check()
```

check_failures



## Explore dependencies

```{r, eval = F}
report1 <- CreatePackageReport(pkg_name = "FACavian")
```



## Build pdf manual
https://stackoverflow.com/questions/16986422/r-create-reference-manual-with-r-cmd-check
R CMD Rd2pdf mypackage


https://stackoverflow.com/questions/30607496/making-an-r-package-pdf-manual-using-devtools
```{r, eval = F}
pack <- "FACavian"
path <- find.package(pack)
system(paste(shQuote(file.path(R.home("bin"), "R")),
    "CMD", "Rd2pdf", shQuote(path)))
```

remove manual
```{r, eval = F}
file.remove("FACavian.pdf")
```


```{r, eval = F}
check(cleanup = FALSE,manual = TRUE,path = getwd())
```



## Keep logs files (?)

```{r, eval = F}
options(keep.source.pkgs = TRUE)
devtools::check()
```




## Rcmd.exe
"Clean and rebuild"
==> Rcmd.exe INSTALL --preclean --no-multiarch --with-keep.source mammalsmilk

"Install restart"
==> Rcmd.exe INSTALL --no-multiarch --with-keep.source mammalsmilk

```{r, eval = F}
devtools::check(env_vars = )
```


devtools::check()

See 'C:/Users/lisanjie/Documents/1_R/git/git-aviary/FACavian.Rcheck/00install.out' for details.
Information on the location(s) of code generating the 'Note's can be
obtained by re-running with environment variable R_KEEP_PKG_SOURCE set
to 'yes'.

