---
title: "z) Building pckgdwn site"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{z) Building pckgdwn site}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## pckgdown functions

### Build packgdwn website

Single function to build whole site:
"This will generate a docs/ directory. The home page will be generated from your package’s README.md, and a function reference will be generated from the documentation in the man/ directory."
```{r, eval = F}
pkgdown::build_site()
```


#### Build piece by piece

pkgdown::clean_site (pkg = ".")

-- Initialising site
```{r, eval = F}
pkgdown::init_site(pkg = ".")
```

Building home 
```{r, eval = F}
pkgdown::build_home(pkg = ".")
```

Building function reference
```{r, eval = F}
pkgdown::build_reference(pkg = ".")
```


Build vignettes.  Note that of 12/15/20 the package ragg is a dependency of pkgdown that doesn't download with pckdown.
```{r, eval = F}
pkgdown::build_articles(pkg = ".")
```


```{r, eval = F}
pkgdown::build_news()
```

